---
title: "LTER data sources"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DT)
library(rhandsontable)
library(readxl)

sources <- read_excel(here::here("data/LTER_data_sources.xlsx"))
```

The following table provides a list of key data sets that may be used to model the effects of global-scale climate forcing on indices of primary productivity across LTER sites. All of the listed LTER Data Sets require the user to adhere to [LTER Data Access Policies](https://lternet.edu/data-access-policy/). Gridded data products acquired from NOAA/OAR/ESRL PSD have no access or use restrictions beyond asking users to cite their [work](https://www.esrl.noaa.gov/psd/data/gridded/). A database specific [Data Use Policy](https://oceaninformatics.ucsd.edu/zoodb/secure/login.php) applies for *ZooDB*, the historical database of zooplankton observations from the California Current Ecosystem. Data from the Northeast Shelf that have been aggregated into the [ecodata](github.com/noaa-edab/ecodata) R package are available for use under a CC0 license.

Double-click into a cell to access its contents. 

```{r, echo = F}
rhandsontable(sources, width = 1000, height = 800) %>%
  hot_cols(colWidths = c(80, 200, 150,150, 100, 100, 150)) %>%
  hot_table(highlightCol = TRUE, highlightRow = TRUE)

# datatable(sources, options = list(
#   pageLength = 100, deferRender = TRUE,
#   scrollY = 400,
#   scroller = TRUE, autowidth = T,fixedColumns = TRUE
# ))

```

