---
title: "Simulating GAMMs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mgcv)
library(mgcViz)
library(nlme)
library(tidyverse)
library(ecodata)
library(forecast)
library(patchwork)
library(magrittr)
library(itsadug)
```

#### Model time series to simulate from
Acquire and process annual primary production data from the Northeast Shelf Large Marine Ecosystem. 

```{r}
ppd <- chl_pp %>% 
  filter(Var %in%
           c("ANNUAL_PPD_MEDIAN SeaWiFS BERENFELD",
             "ANNUAL_PPD_MEDIAN MODIS-Aqua BERENFELD")) 
```

Do the same for chlorophyll *a* data.

```{r}
chl <- chl_pp %>% 
  filter(Var %in%
           c("ANNUAL_CHLOR_A_MEDIAN SeaWiFS PAN",
             "ANNUAL_CHLOR_A_MEDIAN MODIS-Aqua PAN"))
```

We'll also need to pull SST data. The SST data identified below is the average annual SST for the whole Northeast Shelf. It was extracted from the NOAA extended reconstructed sea surface temperature data set (ERSST V5).

```{r}
sst <- ecodata::long_term_sst %>% 
  mutate(id = 1:length(Time))
```


#### Identify time series structures

Using GLS models, we estimated trend and AR(1) correlation structure in time series of chlorophyll *a* and SST in the Northeast Large Marine Ecosystem, and then simulated time series from the fitted models using the `make_sim` function defined below. Simulation models of SST were parameterized with slope terms that varied along the "true" trend estimate of the underlying series. Simulated time series were then fit with GAMMs, where site was included as a random smooth term. 

```{r}

make_sim <- function(df, b0_adjust = 0, b_adjust = 0, n = 21, Site){
  df$id <- 1:length(df$Time)
  site <- unique(df$EPU)
  mod <- gls(Value ~ id, data = df, corr = corAR1(form = ~id))
  phi <- coef(mod$modelStruct$corStruct, unconstrained = F)
  ar.error <- arima.sim(list(order = c(1,0,0),
                          ar = phi),
                     n=n,
                     rand.gen=rnorm,
                     sd = sqrt(var(df$Value)),
                     n.start = NA)
  b1 <- coef(mod)[2]
  b0 <- coef(mod)[1]
  sim <- data.frame(Value = as.numeric(b0 + b0_adjust + 1:n * (b1 + b_adjust) + ar.error),
                    Site = site,
                    Time = as.numeric(1:n))
  return(sim)
}

```

#### Simulate

```{r}
#SST is non-stationary and trend magnitudes vary across latitudes. We generalize simulations of SST by assuming AR1 error structure and varying trend strength, centered around trend derived from NES SST

sstsimlist <- lapply(seq(-0.04, 0.04, length.out = 8), make_sim, df = sst)
sst.df <- do.call(rbind.data.frame, sstsimlist) %>% 
  mutate(Site = factor(rep(paste("Site",letters[1:8]), each = 21)),
         Time = rep(1:21, 8)) %>% 
  dplyr::rename(SST = Value)


#Fit models to PPD time series and simulate
ppd1 <- ppd %>% 
  mutate(Time = as.numeric(as.character(Time))) %>% 
  split(.$EPU) %>% 
  map_df( ~ make_sim(df = .)) %>% 
  mutate(Site = rep(paste("Site",letters[1:4]), each = 21))

ppd2 <- ppd %>% 
  mutate(Time = as.numeric(as.character(Time))) %>% 
  split(.$EPU) %>% 
  map_df( ~ make_sim(df = .)) %>% 
  mutate(Site = rep(paste("Site",letters[5:8]), each = 21))

ppd.sim <- bind_rows(ppd1, ppd2) %>% 
  mutate(Site = factor(Site)) %>% 
  dplyr::rename(PPD = Value)
```

#### Fit GAMMs to examine pressure-response relationship across "ecosystems"

```{r}
pp_merged <- sst.df %>% 
  left_join(.,ppd.sim, by = c("Site","Time"))
  
mixed_pp <- gam(PPD ~ s(SST) + s(SST,Site, bs ='re'),
                  data = pp_merged,
                method = "REML",
                correlation = nlme::corAR1())


par(mfrow = c(1,2),
    mar = c(4,4,4,1))
plot_smooth(mixed_pp, view = "SST", plot_all = "Site",
            col = "#b591c94D", 
            legend_plot_all = NA, 
            hide.label = T, main = "Summed and Partial Effects")
plot_smooth(mixed_pp, view = "SST", 
            plot_all = "Site", col = "#f5d1424D",
            rm.ranef = T, legend_plot_all = NA, 
            hide.label = T,add = T)

plot_smooth(mixed_pp, view = "SST", plot_all = "Site",
            # col = "#b591c94D", 
            # legend_plot_all = NA, 
            hide.label = T, main = "Partial Effects")
# plot_smooth(mixed_pp, view = "SST", 
#             plot_all = "Site", col = "#f5d1424D",
#             rm.ranef = T, legend_plot_all = NA, 
#             hide.label = T,add = T)
```

